An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-06 paper.tex
Introduction

% Paragraph 1: Motivation. At a high level, what is the problem area you
% are working in and why is it important? It is important to set the larger
% context here. Why is the problem of interest and importance to the larger
% community?

@@

One of the main organisations that develops protocol standards is the
Internet Engineering Task Force (IETF). The IETF was founded in 1986,
following on from the US Government-funded effort that developed the early
Internet. It has since grown to become a global community of network
protocol designers, vendors, network operators, and researchers that
develop and publish open network protocol standards and operational
guidelines.  The IETF publishes its standards, and other documents, in
the RFC series.\footnote{\url{https://www.rfc-editor.org}} This series
comprises around 9,000 documents, authored over more than 50 years,
and provides a rich history of the development of the Internet and
its protocols \cite{RFC8700}.

% Paragraph 2: What is the specific problem considered in this paper? This
% paragraph narrows down the topic area of the paper. In the first
% paragraph you have established general context and importance. Here you
% establish specific context and background.

Standardisation is an inherently social and political process
\cite{RFC2026,RFC7282}, requiring cooperation and consensus among a growing
number of stakeholders.

It is important to gain a coherent understanding of the activities that
take place within the IETF, as well as the key factors that may predict
the success of a protocol standard.

% Paragraph 3: "In this paper, we show that...". This is the key paragraph
% in the introduction - you summarize, in one paragraph, what are the main
% contributions of your paper, given the context established in paragraphs
% 1 and 2. What's the general approach taken? Why are the specific results
% significant? The story is not what you did, but rather:
%  - what you show, new ideas, new insights
%  - why interesting, important?
% State your contributions: these drive the entire paper.  Contributions
% should be refutable claims, not vague generic statements.

In this paper, we ...

% Paragraph 4: What are the differences between your work, and what others
% have done? Keep this at a high level, as you can refer to future sections
% where specific details and differences will be given, but it is important
% for the reader to know what is new about this work compared to other work
% in the area.

This paper summarises and expands on our prior work
\cite{mcquistin:2021:characterising,khare:2022:web-we-weave,
healey:2023:power,mcquistin:2023:errare,khare:2023:tracing,
karan:2023:leda,healey:2023:power-frontiers,barnes:2024:temporal}.

% Paragraph 5: "We structure the remainder of this paper as follows." Give
% the reader a road-map for the rest of the paper. Try to avoid redundant
% phrasing, "In Section 2, In section 3, ..., In Section 4, ... ", etc.

We structure the remainder of this paper as follows.
Section \ref{sec:background} introduces the IETF and the Internet Standards
development process, and reviews the data with which we work.
Section \ref{sec:trends-documents} considers trends in document production,
complexity, and correctness.
Section \ref{sec:trends-demographics} moves on to consider the people who
are involved in Internet standards development and how the demographics of
the community have shifted over time.
Section \ref{sec:org-dyn} studies the interaction between participants,
explores who has influence and impact in the IETF and how that influence
is evident in their activities, and the dynamics of communications within
the IETF.
Section \ref{sec:language} considers communications within the IETF from
the perspective of language to explore whether there are differences in
linguistic patterns and language use in more influential participants,
and to consider how the use of language reflects the consensus-driven
nature of the standards development process.
Section \ref{sec:success-factors} then considers factors that lead to
success, exploring what makes a document likely to be published as an
RFC and what makes a successful participant.
Finally, we review related work in Section \ref{sec:related} and conclude
in Section \ref{sec:conclusions}.

>>>>

@@

One of the main organisations that develops protocol standards is the
Internet Engineering Task Force (IETF). The IETF was founded in 1986,
following on from the US Government-funded effort that developed the early
Internet. It has since grown to become a global community of network
protocol designers, vendors, network operators, and researchers that
develop and publish open network protocol standards and operational
guidelines.  The IETF publishes its standards, and other documents, in
the RFC series.\footnote{\url{https://www.rfc-editor.org}} This series
comprises around 9,000 documents, authored over more than 50 years,
and provides a rich history of the development of the Internet and
its protocols \cite{RFC8700}.
This history is further augmented by the rich archive of supporting
documents, work-in-progress drafts, email discussion, and other
metadata about the standards development process that the IETF makes
available.

This gives a unique opportunity to explore the development of the Internet
through its standards. Standards development is an inherently social and
political process \cite{RFC2026,RFC7282}, requiring cooperation and
consensus among a growing number of stakeholders. Due to the importance
of the Internet, it's critical to gain a coherent understanding of the
activities that take place within the IETF, to understand who develops
Internet standards and what factors may predict the success of an idea
or proposal for a new standard.

In this paper, we explore the operation of the IETF to understand who
develops Internet standards, how the participants interact, and what
factors may lead to success.
Section \ref{sec:background} introduces the IETF and the Internet Standards
development process, and reviews the data with which we work.
Section \ref{sec:trends-documents} considers trends in document production,
complexity, and correctness.
item Section \ref{sec:trends-demographics} moves on to consider the people
who are involved in Internet standards development and how the demographics
of the community have shifted over time.
Section \ref{sec:org-dyn} studies the interaction between participants,
explores who has influence and impact in the IETF and how that influence is
evident in their activities, and the dynamics of communications within the
IETF.
Section \ref{sec:language} considers communications within the IETF from
the perspective of language to explore whether there are differences in
linguistic patterns and language use in more influential participants, and
to consider how the use of language reflects the consensus-driven nature of
the standards development process.
Section \ref{sec:success-factors} then considers factors that lead to
success, exploring what makes a document likely to be published as an RFC
and what makes a successful participant.
Finally, we review related work in Section \ref{sec:related} and conclude
in Section \ref{sec:conclusions}.
